---
title: "Summary statistics of acid soil distribution in cropland in SSA"
author: "Sebastian Palmas"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---


This code below estimates the amount of cropland under high acidity. I include general cropland area (based on several cropland areas) and areas under Barley, Maize and Rice and other crops (based on SPAM). It also incoroporates population information under acid soils based in WorldPop rasters. The code returns several summary statistics by country, some graphs and tables that can be used in maps.

This script also prepares the tables used in the acidsoils-explorer Shiny app.

* We use africasoils pH prediction for the continent.
* General cropland area: ESACCI. ESACCI is at 20m, I will decrease the resolution to 1km to facilitate analysis.
* We are using SPAM 2010 v1.1 for the specific crop areas and production
* We use Worldpop for the population estimates


## pH
The pH layer was obtained from ISRIC 250m resolution prediction. 

## ESACCI cropland
## SPAM
These are the physical areas and production several crops at 10 x 10 km grid-cell resolution. This will be the resolution of the analysis and pH and population layers will be warped to match these.

Physical area (A_CROP files): Physical area is measured in a **hectare** and represents the actual area where a crop is grown, not counting how often production was harvested from it. Physical area is calculated for each production system and crop, and the sum of all physical areas of the four production systems constitute the total physical area for that crop. The sum of the physical areas of all crops in a pixel may not be larger than the pixel size.

Production (P_CROP files): Production, for each production system and crop, is calculated by multiplying area harvested with yield. It is measured in **metric tons**. The total production of a crop includes the production of all production systems of that crop.

## WorldPop
The Worldpop has a 1 km resolution and we need to bring it down to 10 km to match the SPAM layer. We need to be careful because the Worldpop layer has population per pixel and the aggregate function needs to be a sum, not a mean of the pixel.

We also need to warp it to match it to the pH layer to be able to make calculations. IF the file already exists, then just read it from the file.

The code above crashes frequently, I will just read the file that it creates instead of creating one everytime the script runs. I need to remember to rerun the code if the original layer changes.



## Steps
The process for this analysis is: 

1. Reading pH and decrease resolution to 1km
2. Warp ESACCI layer to match pH
3. Warp WorldPop to match pH
4. Summarizing data by country
5. Exporting table to use in the acidsoils explorer


## Packages
```{r, message=FALSE}
library(terra)
library(tidyverse)
```

# Reading data
## SSA admin

```{r}
SSA <- vect("F:/Work/Africa/admin/SSA/SSA_admin_simplified.shp")

#Creating table to store results
SSA_names <- as.data.frame(SSA)

#Adding the region
regions <- read_csv("../../data/country_codes_regions.csv")
SSA_names <- inner_join(SSA_names, regions, by=c("country_co", "country_na"))

#adding SSA to use when changing FID2 to ISO3166_1 at the end
SSA_names <- rbind(SSA_names, NA) 
SSA_names$ISO3166_1[45] <- "SSA"
SSA_names$FID2[45] <- sum(SSA_names$FID2, na.rm=TRUE)
```

## SSA tif, pH, croplands, specific crop distributions and population
These rasters are preprocessed in `acidsoils_summary_setup.sh`. These rasters have been resampled to the same 1km resolution and extent.



C3S Values:
10,11,12: Cropland, rainfed
20: Cropland, irrigated or post-flooding
30: Mosaic cropland (>50%)
40: Mosaic cropland (<50%)

Globcover:
11: Post-flooding or irrigated croplands (or aquatic)
14: Rainfed croplands
20: Mosaic cropland (50-70%)
30: Mosaic cropland (20-50%) 


```{r}
#SSA tif
SSA_tif  <- rast("../../data/admin/SSA_admin_simplified_1000m.tif")

#pH
af_PHIHOX <- rast("../../data/afsis/af_PHIHOX_T__M_sd1_1000m_SSA.tif")

#cropland
c3s <- rast("../../data/croplands/C3S-LC-L4-LCCS-Map-1000m-P1Y-2018-v2.1.1_SSA.tif")
globcover <- rast("../../data/croplands/GLOBCOVER_L4_200901_200912_V2.3_1000m_SSA.tif")

#SPAM distribution
A_BARL <- rast("../../data/SPAM/spam2010V1r1_global_A_BARL_A_1000m_SSA.tif")
A_MAIZ <- rast("../../data/SPAM/spam2010V1r1_global_A_MAIZ_A_1000m_SSA.tif")
A_RICE <- rast("../../data/SPAM/spam2010V1r1_global_A_RICE_A_1000m_SSA.tif")

P_BARL <- rast("../../data/SPAM/spam2010V1r1_global_P_BARL_A_1000m_SSA.tif")
P_MAIZ <- rast("../../data/SPAM/spam2010V1r1_global_P_MAIZ_A_1000m_SSA.tif")
P_RICE <- rast("../../data/SPAM/spam2010V1r1_global_P_RICE_A_1000m_SSA.tif")

#WorldPop
AFR_PPP_2020_adj_v2 <- rast("../../data/WorldPop/AFR_PPP_2020_adj_v2_1000m_SSA.tif")
```

## Create table of pixel values
```{r}
SSA_pixel_table <- data.frame(values(SSA_tif))
colnames(SSA_pixel_table) <- "FID2"

SSA_pixel_table$af_PHIHOX <- as.vector(values(af_PHIHOX))

SSA_pixel_table$c3s <- as.vector(values(c3s))
SSA_pixel_table$globcover <- as.vector(values(globcover))

SSA_pixel_table$A_BARL <- as.vector(values(A_BARL))
SSA_pixel_table$A_MAIZ <- as.vector(values(A_MAIZ))
SSA_pixel_table$A_RICE <- as.vector(values(A_RICE))
SSA_pixel_table$P_BARL <- as.vector(values(P_BARL))
SSA_pixel_table$P_MAIZ <- as.vector(values(P_MAIZ))
SSA_pixel_table$P_RICE <- as.vector(values(P_RICE))

SSA_pixel_table$AFR_PPP_2020_adj_v2 <- as.vector(values(AFR_PPP_2020_adj_v2))



SSA_pixel_table <- SSA_pixel_table %>% 
  filter(FID2 != -999) %>% 
  filter(!is.na(af_PHIHOX)) %>% 
  mutate(cropprop = ifelse(c3s == 30, 0.75, ifelse(c3s == 40, 0.25, ifelse(c3s %in% c(10,11), 1, 0))),
         cropprop2 = ifelse(globcover == 20, 0.60, ifelse(globcover == 30, 0.35, ifelse(globcover %in% c(11, 14), 1, 0))),
         ruralpop = AFR_PPP_2020_adj_v2 * cropprop,
         ruralpop2 = AFR_PPP_2020_adj_v2 * cropprop2)
```


#Analysis

## Summarizing total crop areas and rural 


## Summarizing areas of crop by country: Total and acidic

In here we summarize the pixel values by each country and add the results to the final table.
```{r}
SSA_summary_table <- SSA_pixel_table %>%
  group_by(FID2) %>%
  summarise(Crop_AREA = sum(cropprop, na.rm=TRUE),
            Crop_AREA_ACID = sum(cropprop[af_PHIHOX < 56], na.rm=TRUE),
            Crop_AREA2 = sum(cropprop2, na.rm=TRUE),
            Crop_AREA2_ACID = sum(cropprop2[af_PHIHOX < 56], na.rm=TRUE),
            RuralPop = sum(ruralpop, na.rm=TRUE),
            RuralPop_ACID = sum(ruralpop[af_PHIHOX < 56], na.rm=TRUE),
            RuralPop2 = sum(ruralpop2, na.rm=TRUE),
            RuralPop2_ACID = sum(ruralpop2[af_PHIHOX < 56], na.rm=TRUE),
            BARL_AREA = sum(A_BARL),
            BARL_AREA_ACID = sum(A_BARL[af_PHIHOX < 56]),
            MAIZ_AREA = sum(A_MAIZ),
            MAIZ_AREA_ACID = sum(A_MAIZ[af_PHIHOX < 56]),
            RICE_AREA = sum(A_RICE),
            RICE_AREA_ACID = sum(A_RICE[af_PHIHOX < 56]),
            BARL_PROD = sum(P_BARL),
            BARL_PROD_ACID = sum(P_BARL[af_PHIHOX < 56]),
            MAIZ_PROD = sum(P_MAIZ),
            MAIZ_PROD_ACID = sum(P_MAIZ[af_PHIHOX < 56]),
            RICE_PROD = sum(P_RICE),
            RICE_PROD_ACID = sum(P_RICE[af_PHIHOX < 56]))

## Exporting Results
SSA_summary_table <- SSA_names %>% left_join(SSA_summary_table, by="FID2")

write.csv(SSA_summary_table, file="../../output/tables/acid_crops_summary_bycountry.csv", row.names=FALSE)
```

## Summarizing and exporting results for the acidsoils-explorer
In here we summarize the pixel values by each country and add the results to the final table.

```{r}
pH_classes <- c(56, 65, 73, 78)  #original units in ph X 10

#COUNTRY, CROP, TYPE, pH_TYPE, Value
col_names <- c("FID2", "ph_class", "area_km", "prop")
results <- matrix(ncol = length(col_names)) %>% data.frame()
colnames(results) <- col_names

#getting the raster for all
crop_area_results <- SSA_pixel_table %>% 
  group_by(FID2) %>% 
  summarise(ph_class1 = sum(cropprop[af_PHIHOX < pH_classes[1]], na.rm=TRUE),
            ph_class2 = sum(cropprop[af_PHIHOX >= pH_classes[1] & af_PHIHOX < pH_classes[2]], na.rm=TRUE),
            ph_class3 = sum(cropprop[af_PHIHOX >= pH_classes[2] & af_PHIHOX < pH_classes[3]], na.rm=TRUE),
            ph_class4 = sum(cropprop[af_PHIHOX >= pH_classes[3] & af_PHIHOX < pH_classes[4]], na.rm=TRUE),
            ph_class5 = sum(cropprop[af_PHIHOX >= pH_classes[4]], na.rm=TRUE))

#calculating totalSSA
crop_area_results <- rbind(crop_area_results, colSums(crop_area_results))

#calculating prop
crop_area_result_prop <- crop_area_results[,2:6]/rowSums(crop_area_results[,2:6])
crop_area_result_prop <- tibble(FID2 = crop_area_results$FID2, crop_area_result_prop)

#converting both tables to long format and bind them
crop_area_results <- crop_area_results %>% pivot_longer(cols = ph_class1:ph_class5,
                                                        names_to="ph_class",
                                                        values_to="area_km")
crop_area_result_prop <-
  crop_area_result_prop %>% pivot_longer(cols = ph_class1:ph_class5,
                                         names_to="ph_class",
                                         values_to="prop")

crop_area_results <- full_join(crop_area_results, crop_area_result_prop, by=c("FID2", "ph_class"))



#adding lines
results <- rbind(results, crop_area_results)


#change FID2 to ISO3166_1
results$ISO3166_1 <- SSA_names$ISO3166_1[match(results$FID2, SSA_names$FID2)]
results <- results %>% select(ISO3166_1, ph_class, area_km, prop)


write.csv(results, file="C:/Users/S.PALMAS/source/repos/spalmas/acidsoils-explorer/data/cropareas.csv", row.names=FALSE)
```

```{r}
types <- c("Area", "Prop") #c("Area", Prod")
pH_classes <- c(56, 65, 73, 78)  #original units in ph X 10
crops <- c("BARL", "MAIZ", "RICE")

#COUNTRY, CROP, TYPE, pH_TYPE, Value
col_names <- c("FID2", "ph_class", "area_km", "prop", "crop", "type", "unit")
results <- matrix(ncol = length(col_names)) %>% data.frame()
colnames(results) <- col_names

for (crop in crops){
  #crop <- crops[1]  #to test
  for (type in types){ 
    #type <- types[1]
    if (type == "Area"){
      type_code <- "A_"
      type_unit <- "km2"
    } else if (type == "Prod"){
      type_code <- "P_"
      type_unit <- "MT"
    }
    
    
    SSA_pixel_table$type_crop <- SSA_pixel_table[,paste0(type_code, crop)]
    
    #getting the raster for all
    crop_type_results <- SSA_pixel_table %>% 
      group_by(FID2) %>% 
      summarise(ph_class1 = sum(type_crop[af_PHIHOX < pH_classes[1]], na.rm=TRUE),
                ph_class2 = sum(type_crop[af_PHIHOX >= pH_classes[1] & af_PHIHOX < pH_classes[2]], na.rm=TRUE),
                ph_class3 = sum(type_crop[af_PHIHOX >= pH_classes[2] & af_PHIHOX < pH_classes[3]], na.rm=TRUE),
                ph_class4 = sum(type_crop[af_PHIHOX >= pH_classes[3] & af_PHIHOX < pH_classes[4]], na.rm=TRUE),
                ph_class5 = sum(type_crop[af_PHIHOX >= pH_classes[4]], na.rm=TRUE))
    
    #calculating totalSSA
    crop_type_results <- rbind(crop_type_results, colSums(crop_type_results))
    
    #calculating prop
    crop_type_result_prop <- crop_type_results[,2:6]/rowSums(crop_type_results[,2:6])
    crop_type_result_prop <- tibble(FID2 = crop_type_results$FID2, crop_type_result_prop)

    #converting both tables to long format and bind them
    crop_type_results <- crop_type_results %>% pivot_longer(cols = ph_class1:ph_class5,
                                                          names_to="ph_class",
                                                          values_to="area_km")
    crop_type_result_prop <-
      crop_type_result_prop %>% pivot_longer(cols = ph_class1:ph_class5,
                                                                    names_to="ph_class",
                                                                    values_to="prop")
    
    crop_type_results <- full_join(crop_type_results, crop_type_result_prop, by=c("FID2", "ph_class"))
    
    
    #fill results table
    crop_type_results$crop <- crop
    crop_type_results$type <- type
    crop_type_results$unit <- type_unit
    
    #adding lines
    results <- rbind(results, crop_type_results)
  }
}


#change FID2 to ISO3166_1
results$ISO3166_1 <- SSA_names$ISO3166_1[match(results$FID2, SSA_names$FID2)]
results <- results %>% select(ISO3166_1, ph_class, area_km, prop, crop, type, unit)


write.csv(results, file="C:/Users/S.PALMAS/source/repos/spalmas/acidsoils-explorer/data/by_crop_areas.csv", row.names=FALSE)
```

Population table

```{r}
pH_classes <- c(56, 65, 73, 78)  #original units in ph X 10

#COUNTRY, CROP, TYPE, pH_TYPE, Value
col_names <- c("FID2", "ph_class", "population", "prop")
pop_results <- matrix(ncol = length(col_names)) %>% data.frame()
colnames(pop_results) <- col_names

#getting the raster for all
pop_results <- SSA_pixel_table %>% group_by(FID2) %>% 
  summarise(ph_class1 = sum(ruralpop[af_PHIHOX < pH_classes[1]], na.rm=TRUE),
            ph_class2 = sum(ruralpop[af_PHIHOX >= pH_classes[1] & af_PHIHOX < pH_classes[2]], na.rm=TRUE),
            ph_class3 = sum(ruralpop[af_PHIHOX >= pH_classes[2] & af_PHIHOX < pH_classes[3]], na.rm=TRUE),
            ph_class4 = sum(ruralpop[af_PHIHOX >= pH_classes[3] & af_PHIHOX < pH_classes[4]], na.rm=TRUE),
            ph_class5 = sum(ruralpop[af_PHIHOX >= pH_classes[4]], na.rm=TRUE))


#calculating prop
pop_results_prop <- pop_results[,2:6]/rowSums(pop_results[,2:6])
pop_results_prop <- tibble(FID2 = pop_results$FID2, pop_results_prop)

#converting both tables to long format and bind them
pop_results <- pop_results %>% pivot_longer(cols = ph_class1:ph_class5,
                                                      names_to="ph_class",
                                                      values_to="population")
pop_results_prop <-
  pop_results_prop %>% pivot_longer(cols = ph_class1:ph_class5,
                                         names_to="ph_class",
                                         values_to="prop")

pop_results <- full_join(pop_results, pop_results_prop, by=c("FID2", "ph_class"))

#change FID2 to ISO3166_1
pop_results$ISO3166_1 <- SSA_names$ISO3166_1[match(pop_results$FID2, SSA_names$FID2)]
pop_results <- pop_results %>% select(ISO3166_1, ph_class, population, prop)


write.csv(pop_results, file="C:/Users/S.PALMAS/source/repos/spalmas/acidsoils-explorer/data/population.csv", row.names=FALSE)
```