---
title: "Estimating acid soil distribution in cropland at the continental scale."
author: "Sebastian Palmas"
date: "2019-07-27"
output: pdf_document
---

```{r packages, message=FALSE}
library(countrycode)
library(ggplot2)
library(tidyverse)
```


```{r, engine = 'bash', echo = F, eval = FALSE}

#7z l ~/Downloads/acidsoils-20190723T082159Z-001.zip  #list files
#7z x -aoa	~/Downloads/acidsoils-20190723T082159Z-001.zip Data/
7z x -aoa	~/Downloads/acidsoils-20190723T082159Z-001.zip Data/

```


# Acid soils

With the help of [Google Earth Engine (GEE)](https://earthengine.google.com/) we can analyze any continental spatial data in seconds.

This code below estiamtes the amount of cropland with an estimated high acidity (pH < 5.6). It then summarises the results by country.

* We are using [CCI LAND COVER - S2 PROTOTYPE LAND COVER 20M MAP OF AFRICA 2016Copernicus Global Land Cover](http://2016africalandcover20m.esrin.esa.int/).

* We use africasoils pH prediction for the continent.

## LULC
```{r LULC, message=FALSE}
#I also remove Cabo Verde since it is not included in the LULC data
#to solve some issues, the Koualou Area (UU), Halaib Triangle (SU) and the Halaib Triangle (UU) are ommited in the results
#countries_to_omit<- c('CV', 'UU')

json <- jsonlite::fromJSON('../Data/GeoJSON/SSA_LULC_Histogram.geojson')
LULC <- json$features$properties$histogram %>%  `/` (1000000) %>% `*` (20^2)   #Original data in pixels (of 20^2 m2), *20^2 to convert to m2, and /1000000 to convert to km2
LULC <- cbind(json$features$properties$country_co, json$features$properties$country_na,  LULC) %>% 
  filter(!country_co %in% countries_to_omit)

#LULC_classes <- read_delim('../Data/ESACCI-LC-L4-LC10-Map-20m-P1Y-2016-v1.0/ESACCI-LC_S2_Prototype_ColorLegend.csv', delim = ';')
colnames(LULC) <- c('country_co','country_na', 'Tree cover areas','Open Water',
                    'Shrubs cover areas','NA','Grassland','Cropland',
                    'Vegetation aquatic or regularly flooded', 'Lichens Mosses / Sparse vegetation','Bare areas','Built up areas')

write_csv(LULC[,-6], path = '../Results/LULC.csv')

```

## Crop areas by pH classes
This makes a table of the crop areas by pH classes by country. Its in long format for easy use in ggplot later
```{r }
#Croland area under different ph classes 
json <- jsonlite::fromJSON('../Data/GeoJSON/ph_class_bands_table.geojson')
cropareas <- json$features$properties %>% 
  select(country_co, ph_class0, ph_class1, ph_class2, ph_class3, ph_class4)

#calculating SSA total
SSAsum <- cropareas[,2:6] %>% colSums()  
cropareas <- cropareas %>% rbind(c(NA, SSAsum)) #adding values to table
cropareas$country_co[nrow(cropareas)] <- "SSA"  #adding SSA to country_co

cropareas <- cropareas %>% 
  gather(key = "ph_class", value = "area_km", -country_co) %>% 
  mutate(ph_class = factor(ph_class),
    area_km = round(area_km/1000000))
str(cropareas)
#save table to rds to use for shiny app
saveRDS(cropareas, 'C:/Users/S.PALMAS/source/repos/spalmas/acidsoils-explorer/data/cropareas.rds')
write.csv(cropareas, 'D:/OneDrive - CGIAR/AcidSoils/Results/cropareas.csv')

```

## Population analysis
```{r population}
json <- jsonlite::fromJSON('../Data/GeoJSON/pop_class_bands_table.geojson' )
pop <- json$features$properties %>% 
  select(country_co, SSApop, pop_class0, pop_class1, pop_class2, pop_class3, pop_class4) 

#calculating SSA total
SSAsum <- pop[,2:7] %>% colSums()  
pop <- pop %>% rbind(c(NA, SSAsum)) #adding values to table
pop$country_co[nrow(pop)] <- "SSA"  #adding SSA to country_co

#population table for Cristina
popcsv <- pop %>%
  mutate(pop2020 = round(SSApop),
         ruralpop2020 = round(pop_class0 + pop_class1 + pop_class2 + pop_class3 + pop_class4),
         country.name = countrycode(country_co, origin = "fips", destination = "country.name")) %>% 
  select(country_co, country.name, pop2020, ruralpop2020)
popcsv$country.name[popcsv$country_co == "OD"] <- "South Sudan"
popcsv$country.name[popcsv$country_co == "SSA"] <- "SubSaharan Africa"
popcsv[1:(nrow(popcsv)-1),] <- popcsv[popcsv$country.name[1:(nrow(popcsv)-1)] %>% order,] #reorder by country names

write_csv(popcsv, "D:/OneDrive - CGIAR/AcidSoils/Results/population.csv")

pop <- pop %>% 
  gather(key = "ph_class", value = "population", -country_co) %>% 
  mutate(ph_class = factor(ph_class))
str(pop)
#save table to rds to use for shiny app
saveRDS(pop, 'C:/Users/S.PALMAS/source/repos/spalmas/acidsoils-explorer/data/population.rds')



```

## Computing some new columns
```{r newcolumns}
#export values to a csv
acidsoils$propacidcroparea[is.nan(acidsoils$propacidcroparea)] <- NA
write_csv(acidsoils, '../Results/acidsoils_bycountry.csv')
head(acidsoils)

```


## Countries arranged by the estimated cropland area under acid soils 
```{r plot1, echo = FALSE, message=FALSE}
p2 <- ggplot(acidsoils %>% filter(acidcroparea_km > 0), aes(x = reorder(country_na, -acidcroparea_km), y = acidcroparea_km)) +
  geom_bar(stat = "identity") +
  ylab('Cropland area with high acidity (km2)') +
  theme(axis.text.x = element_text(angle = 90))+
  xlab('Country')
p2
```

## Percentage of country cropland under high acidiry
```{r perc, echo = FALSE}
p2 <- ggplot(acidsoils %>% filter(propacidcroparea > 0), aes(x = reorder(country_na, -propacidcroparea), y = propacidcroparea)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90))+
  ylab('Percertage of country cropland high acidity (%)') + xlab('Country')
p2
```

## Rural Population living on acid soils
```{r perc, echo = FALSE}
p2 <- ggplot(acidsoils %>% filter(pop_acidcropland > 0), aes(x = reorder(country_na, -pop_acidcropland), y = pop_acidcropland)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))+
  ylab('Population living on acid soils') + xlab('Country')
p2
```

# Applications
* Estimate the amount of inputs required to increase pH to a certain level.
* With data on limestone production locations, we can estimate the cost of taking it to the farms,
* Make an analysis of prioritization of areas to increase yields.
