---
title: "Setup for the Ensemble Prediction of Market Maize Prices in Ethiopia"
author: "Sebastian Palmas"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

# Introduction
Setup for the ensemble prediction of maize market prices in Ethiopia. It creates tables for calibration and validation.

## Packages
We use the "sp", "raster" and "rgdal" R packages for creating and manipulating spatial data in this exercise. They are all add-on packages that have to be installed from CRAN.
```{r, message = FALSE}
rm(list=ls()) 
library(dplyr)
library(rgdal)
library(terra)
```


# I. PREPARING CSA DATA
This data is made up of maize prices reported by CSA across Ethiopia in the 2014-2019. 

This is the CSA raw data.

```{r}
CSA <- read.csv("F:/Work/Ethiopia/CSA_RetailPrice/CSA_RetailPrice_BarleyMaizeWheat_2013_19.csv")
```

## 1) Cleaning the data
### Changing factors to numbers
```{r}
#Data has points instead of NAs
CSA[CSA == "."] <- NA

#Removing points without coordinates
CSA <- CSA[!is.na(CSA$Longitude) & !is.na(CSA$Latitude),]

#converting factor to numeric
CSA$Longitude <- as.numeric(CSA$Longitude)
CSA$Latitude <- as.numeric(CSA$Latitude)
CSA$barley_unmilled_White <- as.numeric(CSA$barley_unmilled_White)
CSA$maize_Unmilled <- as.numeric(CSA$maize_Unmilled)
CSA$wheat_unmilled_White <- as.numeric(CSA$wheat_unmilled_White)
```



### Creating a date column
```{r}
months.name <- c("january", "february", "march", "april", "may", "june", "july", "august", "september", "october", "november", "december")

#creating month code number based on list above
CSA$Month_code <- lubridate::month(match(CSA$Month, months.name))

#creating column in date format
CSA$Date <- as.Date(with(CSA, paste(Year, Month_code, 1,sep="-")), "%Y-%m-%d")
```

## 2) Correcting temporal trend

We eliminate the temporal trend and scale it back by doing a z-transformation using the mean and sd of each year and month.
```{r}
CSA_summary_price <- CSA %>% 
  group_by(Year, Month_code) %>% 
  summarise(barley_unmilled_White_mean = mean(barley_unmilled_White, na.rm=TRUE),
            maize_Unmilled_mean = mean(maize_Unmilled, na.rm=TRUE),
            wheat_unmilled_White_mean = mean(wheat_unmilled_White, na.rm=TRUE),
            #Standard deviation
            barley_unmilled_White_sd = sd(barley_unmilled_White, na.rm=TRUE),
            maize_Unmilled_sd = sd(maize_Unmilled, na.rm=TRUE),
            wheat_unmilled_White_sd = sd(wheat_unmilled_White, na.rm=TRUE))

#joining the data to the CSA
CSA <- left_join(CSA, CSA_summary_price, by = c("Year", "Month_code"))

#Z-score standardization
CSA <- CSA %>%
  mutate(barley_unmilled_White_norm = (barley_unmilled_White - barley_unmilled_White_mean) / barley_unmilled_White_sd,
         maize_Unmilled_norm = (maize_Unmilled - maize_Unmilled_mean) / maize_Unmilled_sd,
         wheat_unmilled_White_norm = (wheat_unmilled_White - wheat_unmilled_White_mean) / wheat_unmilled_White_sd)

```

## 3) Adding previous month price as column
We need to match the previous month prices from each market to each row.

Because of the many rules, I'll do this in a for loop even though it is slow.

```{r}
#column to store values
CSA$barley_unmilled_White_norm_t0 <- NA
CSA$maize_Unmilled_norm_t0 <- NA
CSA$wheat_unmilled_White_norm_t0 <- NA

for (r in 1:nrow(CSA)){
  #r <- 1 #to test
  Market1 <- CSA$Market[r]
  
  Month_code1 <- CSA$Month_code[r]
  Year1 <- CSA$Year[r]
  
  if(Month_code1 == 1){
    Month_code0 <- 12
    Year0 <- Year1 - 1
  } else {
    Month_code0 <- Month_code0 - 1
    Year0 <- Year1
  }
  
  CSA0 <- CSA %>% filter(Market == Market1 & Month_code == Month_code0 & Year == Year0)
  
  if(nrow(CSA0) != 0){
    CSA$barley_unmilled_White_norm_t0[r] <- CSA0$barley_unmilled_White_norm
    CSA$maize_Unmilled_norm_t0[r] <- CSA0$maize_Unmilled_norm
    CSA$wheat_unmilled_White_norm_t0[r] <- CSA0$wheat_unmilled_White_norm
  }
}
```

## 4) Creating spatial object from the CSA data

Now that we have a dataset will full coordinates and prices, I'll create a shapefile to extract the values from the rasterstack.

```{r}
#creating spatial object with those coordinates
CSA_vect <- vect(cbind(CSA$Longitude, CSA$Latitude), atts=CSA)

# wgs84.prj: projection for coordinates in prices csv
crs(CSA_vect) <- "+proj=longlat +datum=WGS84 +no_defs"
```

The raster files in the Markus stack use [Lambert azimuthal equal-area projection](https://en.wikipedia.org/wiki/Lambert_azimuthal_equal-area_projection), which better preserves areas and distances compared with the WGS84 projection. We transform the projection of the maize price shapefile to match this projection since it must share the same projection system with the rasters in any spatial process.

```{r}
#reading PH to use its projection
PH <- rast("F:/Work/MarkusStacks/ET_250m_2019/ET_500m_2019_resampled/PH.tif")

#Projecting the vector to this projection
CSA_vect <- terra::project(CSA_vect, crs(PH)[1])

#export vector to use for predictions
#writeVector(CSA_vect,
#            filename = "../../data/prices/CSA_market_maize_prices_2016.shp",
#            overwrite=TRUE)
```

Changing the coordinates in the table to the correct projection
```{r}
CSA$Longitude <- geom(CSA_vect)$x
CSA$Latitude <- geom(CSA_vect)$y

```


```{r, echo=FALSE}
ETH_adm <- vect("F:/Work/GADM/gadm36_levels_shp/gadm36_ETH_shp/gadm36_ETH_0.shp") %>% project(crs(PH)[1])
plot(CSA_vect, pch = 20, col = "Red", main = "Locations With Known CSA Prices")
lines(ETH_adm)
```


# II. PREPARING TIME-INVARIANT RASTERSTACK
There are multiple spatial covariates. Meher/Belg seasonality, weather, soil and amenities. Also we add latitude and longitude data.

## 1) Meher/Belg location
Raster version of the Meher/Belg distribution. This original vector version comes from. The rasterized version is prepared in `prepare_rasters.sh` file.
```{r}
Meher_Belg <- rast("../../data/Meher_Belg_Regions/Meher_Belg_Regions.tif")

CSA <- cbind(CSA,
             terra::extract(Meher_Belg, CSA_vect,drop=TRUE))
```


```{r, echo=FALSE}
plot(Meher_Belg)
```

## 2) Add Latitude and Longitude Rasters
We add longitude and latitude rasters to the stack so that our model considers location in its prediction. The longitude and latitude rasters will have the longitude values and latitude values respectively as the cell values. We use one raster from the stack created earlier as a reference, to ensure the latitude and longitude rasters have the same properties.

Create the longitude and latitude rasters in 5 steps: convert the Tanzania's "level 0" shapefile -downloaded earlier- projection to the Lambert azimuthal equal-area projection; transform it to a raster with the reference raster's properties; extract the longitude and latitude values from the created raster; assign the longitude and latitude values to two copies of the created raster; name the raster cell values and add them to the rasters stack.

This most likely can be done more efficiently with r.latlong in GRASS, but I couldn't make it work.

```{r, eval=FALSE}
#Extacting longitude and latitude values for each column and row
longitudes <- xFromCol(rasterstack[[1]], 1:dim(rasterstack[[1]])[2]) 
longitudes <- rep(longitudes, times = dim(rasterstack[[1]])[1])
latitudes <- yFromRow(rasterstack[[1]], 1:dim(rasterstack[[1]])[1])
latitudes <- rep(latitudes, times = dim(rasterstack[[1]])[2])
latitudes <- sort(latitudes, decreasing=TRUE)

#Creating the rasters and assigning the values
natbnd.raster.long <- rasterstack[[1]]
values(natbnd.raster.long) <- longitudes
names(natbnd.raster.long) <- "Longitude"  #changing the name of the values
writeRaster(natbnd.raster.long, filename = "F:/Work/MarkusStacks/ET_250m_2019/ET_500m_2019_resampled/long.tif", overwrite=TRUE)
natbnd.raster.lat <- rasterstack[[1]]
values(natbnd.raster.lat) <- latitudes
names(natbnd.raster.lat) <- "Latitude"  #changing the name of the values
writeRaster(natbnd.raster.lat, filename = "F:/Work/MarkusStacks/ET_250m_2019/ET_500m_2019_resampled/lat.tif", overwrite=TRUE)
```

## 3) Raster Data on Weather, Soil and Amenities
The raster files contain data on weather elements, soil properties and distance to infrastructure/amenities. In this exercise, these files are in a sub-directory of the working directory. You can download them [here](https://osf.io/j8y3z/). The table below gives a short description of each file and a link to the source website.

|Raster | Short description | Derived from:
|:-----|:---------|:-----|
|BIO1 | Mean annual temperature (째C * 10, 1979-2013) | [CHELSA](http://chelsa-climate.org/downloads/)
|BIO7 | Mean annual temperature range (째C * 10, 1979-2013) | [CHELSA](http://chelsa-climate.org/downloads/)
|BIO12 | Mean annual precipitation (mm/yr, 1979-2013) | [CHELSA](http://chelsa-climate.org/downloads/)
|BIO15 | Mean rainfall seasonality (cv, 1979-2013) | [CHELSA](http://chelsa-climate.org/downloads/)
|BPP15 | Predicted 2015 building presence probability | [Africagrids.net](ftp://africagrids.net/)
|CEC | Predicted topsoil cation exchange capacity (cmol/kg) | [SoilGrids](https://soilgrids.org)
|CPP15 | Predicted 2015 cropland presence probability | [Africagrids.net](ftp://africagrids.net/)
|DCELL | Distance to Cell Tower or WiFi point (km) | [UnwiredLabs](https://unwiredlabs.com)
|DFRES | Distance to forest reserves (km) | [Protected Planet](https://www.protectedplanet.net/)
|DGRES | Distance to game reserves (km) | [Protected Planet](https://www.protectedplanet.net/)
|DGRID | Distance to X (km) | [X](ftp://africagrids.net/)
|DHRSL | Distance to High-Resolution Settlement Layer (km) | [CIESIN](https://ciesin.columbia.edu/data/hrsl/)
|DNLT | Distance to BlackMarble night-lights (km) | [NASA](https://earthobservatory.nasa.gov/features/NightLights/page3.php)
|DOR1 | Distance to main roads (km) | [Geofabrik](#https://www.geofabrik.de/data/download.html)
|DOR2 | Distance to any known road (km) | [Geofabrik](#https://www.geofabrik.de/data/download.html)
|DOWS | Distance to inland water bodies (km) | [SurfaceWater](#https://global-surface-water.appspot.com/)
|DPARK | Distance to national parks & conservation areas (km) | [Protected Planet](https://www.protectedplanet.net/)
|DPOP1 | Distance to city or town centers (km) | [Geofabrik](#https://www.geofabrik.de/data/download.html)
|DPOP2 | Distance to village or hamlet centers (km) | [Geofabrik](#https://www.geofabrik.de/data/download.html)
|EVI | Average enhanced vegetation index (2000-2016) | [Africagrids.net](ftp://africagrids.net/)
|FIRE | Detected fire density (2001-2015) | [Africagrids.net](ftp://africagrids.net/)
|GBD | Building footprint densities (2017) | [DigitalGlobe](https://explore.digitalglobe.com/Tanzania-Building-Footprints.html)
|GCCP | USGS cropland extent prediction (2015) | [USGS](https://croplands.org/app/map?lat=0.17578&lng=-0.17578125&zoom=2)
|GFPL | Global floodplain extent (2017) | [USGS](https://www.usgs.gov/centers/eros/science/usgs-eros-archive-landsat-landsat-level-3-dynamic-surface-water-extent-dswe?qt-science_center_objects=0#qt-science_center_objects)
|LCB | Bare /sparse vegetation cover fraction (2015) | [Copernicus](http://lcviewer.vito.be/)
|LCC | Cropland cover fraction (2015) | [Copernicus](http://lcviewer.vito.be/)
|LCS | Shrubland cover fraction (2015) | [Copernicus](http://lcviewer.vito.be/)
|LCT | Forest cover fraction (2015) | [Copernicus](http://lcviewer.vito.be/)
|LCU | Built-up cover fraction (2015) | [Copernicus](http://lcviewer.vito.be/)
|LSTD | Average day-time land surface temp. (째C , 2001-2017) | [Africagrids.net](ftp://africagrids.net/)
|LSTN | Average night-time land surface temp. (째C , 2001-2017) | [Africagrids.net](ftp://africagrids.net/)
|MB1 | Average MOD13Q1 band 1 reflectance (2001-2016)  | [Africagrids.net](ftp://africagrids.net/)
|MB2  | Average MOD13Q1 band 2 reflectance (2001-2016) | [Africagrids.net](ftp://africagrids.net/)
|MB3 | Average MOD13Q1 band 3 reflectance (2001-2016) | [Africagrids.net](ftp://africagrids.net/)
|MB7 | Average MOD13Q1 band 7 reflectance (2001-2016) | [Africagrids.net](ftp://africagrids.net/)
|MDEM | Elevation above mean sea level (m) | [MERITDEM](http://hydro.iis.u-tokyo.ac.jp/~yamadai/MERIT_DEM/)
|NPPA | Average annual net primary productivity (kg/m2, 2000-2015) | [Africagrids.net](ftp://africagrids.net/)
|NPPS | Variance annual net primary productivity (2000-2015) | [Africagrids.net](ftp://africagrids.net/)
|PARA | Average fAPAR (2000-2017) | [Africagrids.net](ftp://africagrids.net/)
|PARV | Variance fAPAR (2000-2017) | [Africagrids.net](ftp://africagrids.net/)
|PH | Predicted topsoil pH | [SoilGrids](https://soilgrids.org)
|S1VV | Average Sentinel 1 VV radar backscatter (2016) | [Africagrids.net](ftp://africagrids.net/)
|S2B11 | Average Sentinel 2 SWIR1 reflectance (2016-2017) | [Africagrids.net](ftp://africagrids.net/)
|S2B12 | Average Sentinel 2 SWIR2 reflectance (2016-2017) | [Africagrids.net](ftp://africagrids.net/)
|SLOPE | Slope (%) | [MERITDEM](http://hydro.iis.u-tokyo.ac.jp/~yamadai/MERIT_DEM/)
|SND | Predicted topsoil sand content (%) | [SoilGrids](https://soilgrids.org)
|SOC | Predicted topsoil organic carbon content (g/kg) | [SoilGrids](https://soilgrids.org)
|TIM | Topographic index map | [MERITDEM](http://hydro.iis.u-tokyo.ac.jp/~yamadai/MERIT_DEM/)
|WPOP | Population density prediction (2015) | [WorldPop](#http://www.worldpop.org.uk/data/get_data/)
|WPP17 | Population density prediction (2015) | [WorldPop](#http://www.worldpop.org.uk/data/get_data/)

We create a raster object with all these layers. Some Markus files had slighly different projection and extent, I corrected these files in `prepare_rasters.sh`.

```{r}
rasterlist <- list.files(path = "F:/Work/MarkusStacks/ET_250m_2019/ET_500m_2019_resampled/", 
                         pattern = "*.tif$", 
                         full.names = TRUE) # Character vector of relative filepaths

#eliminate lat long rasters because coordinates where already in CSA dataset
lat_raster <- grep("lat.tif", rasterlist)
rasterlist <- rasterlist[-lat_raster]

long_raster <- grep("long.tif", rasterlist)
rasterlist <- rasterlist[-long_raster]

#creating rasterstack
rasterstack <- rast(rasterlist)

#extracting values within a raster
CSA <- cbind(CSA,
             terra::extract(rasterstack, CSA_vect, 
                            #buffer=5000, # Meters
                            drop=TRUE))
```






# III. ADDING SEASONAL RAINFALL DATA FROM

## Specific rainfall seasons

These are the rules for assigning rainfall estiamtes to each row:

* Meher (Kiremt) rainfall that runs from March to September are assigned to rows between September to August.
* In Belg-receiving areas, Belg rains between February to May are assigned to prices between June and August.

This code below adds the seasonal rainfalls depending on the year and month of each row.

```{r}
#Rainfall seasons to analyze
kiremt_rainfall_seas <- c("06", "07", "08") #June to August
belg_rainfall_seas <- c("02", "03", "04", "05") #february to may 

#list of seasons names to use in the loop
rainfall_seasons <- c("kiremt_rainfall_seas", "belg_rainfall_seas")

#dekads to use
dekads <- c(1,2,3)

#years to extract
years <- CSA$Year %>% unique() %>% sort()

#table to store results to later cbind to CSA
results <- rep(NA, times = nrow(CSA))

#chirps dataset is in a different projection
CSA_vect_forchirps <- CSA_vect
CSA_vect_forchirps <- project(CSA_vect_forchirps, "+proj=longlat +datum=WGS84 +no_defs")

for(rainfall_season in rainfall_seasons){
  #rainfall_season <- rainfall_seasons[1]  #to test
  for (year in years){
    #year <- 2013 #to test
    
    #tifs of all combinations of all dekads and months for that season
    rainfall_season_rasters <- rast(apply(expand.grid("F:/Work/CHIRPS/ftp.chg.ucsb.edu/pub/org/chg/products/CHIRPS-2.0/africa_dekad/tifs/chirps-v2.0.",
                                                      year, ".", get(rainfall_season), ".", dekads, ".tif"), 1, paste, collapse=""))
    
    #Getting the sums of rainfall for all points
    values <- rowSums(terra::extract(rainfall_season_rasters, CSA_vect_forchirps, drop=TRUE))
    
    #Depending on the season, we store the prices in specific months
    if(rainfall_season == "kiremt_rainfall_seas"){
      #months to store (all of them for Kiremt)
      season_price_months1 <- c(9,10,11,12)
      season_price_months2 <- c(1,2,3,4,5,6,7,8)
    
      #which months to store (kiremt prices go to the next year)
      rows1 <- (CSA$Year==year) & (CSA$Month_code %in% season_price_months1)
      rows2 <- (CSA$Year==year+1) & (CSA$Month_code %in% season_price_months2)
      
      #store only those in the selected rows
      results[rows1|rows2]<- values[rows1|rows2] #the sum of the rainfall for dekads
    
    } else if(rainfall_season == "belg_rainfall_seas") {
      #months to store (all of them for Kiremt)
      season_price_months1 <- c(6,7,8) 
      
      #which months to store (kiremt prices go to the next year)
      rows1 <- (CSA$Year==year) & (CSA$Month_code %in% season_price_months1) & (CSA$Meher_Belg_Regions == 2)
      
      #store only those in the selected rows
      results[rows1]<- values[rows1] #the sum of the rainfall for dekads
    }
  }
}

#bind the results to a column
CSA$rainfall_season <- results
```

## All columns for all seasons
I'll add multiple rainfall seasons summaries to the rows. This code below adds the seasonal rainfalls depending on the year of the row.

```{r, eval=FALSE}
#Rainfall seasons to analyze
kiremt_rainfall_seas <- c("06", "07", "08") #June to August
belg_rainfall_seas <- c("02", "03", "04", "05") #february to may 
meher_barley_rainfall_seas <- c("05", "06", "07", "08","09", "10") #may to october
meher_maize_rainfall_seas <- c("03", "04", "05", "06", "07", "08", "09") #march to september
meher_wheat_rainfall_seas <- c("05", "06", "07", "08","09", "10") #march to september

#list of seasons names to use in the loop
rainfall_seasons <- c("kiremt_rainfall_seas", "belg_rainfall_seas", "meher_barley_rainfall_seas", "meher_maize_rainfall_seas", "meher_wheat_rainfall_seas")

#dekads to use
dekads <- c(1,2,3)

#years to extract
years <- CSA$Year %>% unique() %>% sort()

#table to store results to later cbind to CSA
results <- data.frame(matrix(ncol = length(rainfall_seasons), nrow = nrow(CSA)))
colnames(results) <- c(rainfall_seasons)

#chirps dataset is in a different projection
CSA_vect_forchirps <- CSA_vect
CSA_vect_forchirps <- project(CSA_vect_forchirps, "+proj=longlat +datum=WGS84 +no_defs")

for(rainfall_season in rainfall_seasons){
  #rainfall_season <- rainfall_seasons[1]  #to test
  for (year in years){
    #year <- 2013 #to test
    
    #tifs of all combinations of all dekads and months for that season
    rainfall_season_rasters <- rast(apply(expand.grid("F:/Work/CHIRPS/ftp.chg.ucsb.edu/pub/org/chg/products/CHIRPS-2.0/africa_dekad/tifs/chirps-v2.0.",
                                                      year, ".", get(rainfall_season), ".", dekads, ".tif"), 1, paste, collapse=""))
    
    #extracting for all dekads and taking the sum. Only store those values that match the year
    results[CSA$Year==year,rainfall_season]<- rowSums(terra::extract(rainfall_season_rasters, CSA_vect_forchirps, drop=TRUE))[CSA$Year==year] #the sum of the rainfall for dekads
    
  }
}
#cbind the results to CSA table that will be exportable
CSA <- cbind(CSA, results)
```

## Lagged rainfall season

Rainfall of the last months

```{r}
#List of months of lag that should be extracted (rainfall in the previous n months)
lags <- c(3,6)

#table to store results to later cbind to CSA
results <- data.frame(matrix(ncol = length(lags), nrow = nrow(CSA)))
colnames(results) <- paste0("lagrainfall", lags)

for(l in lags){
  #l <- lags[1]
  for (year in years){
    #year <- 2013 #to test
    for(month in months.name){
      #month <- "january"
      month.code <- match(month, months.name)  #initial month code
      
      month.lag.code <- (month.code-l):(month.code-1)
      month.lag.code[month.lag.code<=0] <-  month.lag.code[month.lag.code<=0] + 12
      
      #adding a leading 0 to single digits
      month.lag.code <- sprintf("%02d", month.lag.code)
      
      #tifs of all combinations of all dekads and months for that season
      rainfall_season_rasters <- rast(apply(expand.grid("F:/Work/CHIRPS/ftp.chg.ucsb.edu/pub/org/chg/products/CHIRPS-2.0/africa_dekad/tifs/chirps-v2.0.",
                                                        year, ".", month.lag.code, ".", dekads, ".tif"), 1, paste, collapse=""))
      
      #extracting for all dekads and taking the sum. Only store those values that match the year
      results[CSA$Year==year & CSA$Month==month,paste0("lagrainfall", l)]<- rowSums(terra::extract(rainfall_season_rasters, CSA_vect_forchirps, drop=TRUE))[CSA$Year==year & CSA$Month==month] #the sum of the rainfall for dekads
      
    }
    
  }
}

#cbind the results to CSA table that will be exportable
CSA <- cbind(CSA, results)
```

# IV. SELECTING VARIABLES TO USE IN THE PREDICTION
```{r}
CSA_barley <- CSA %>%
  filter(!is.na(barley_unmilled_White_norm)) %>%
  dplyr::select(barley_unmilled_White_norm,
                barley_unmilled_White_norm_t0,
                Month_code, Longitude, Latitude,
                BIO1, BIO12, BIO15, BIO7, CEC,
                DCELL, DGRID, DNFPA, DOR1, DOR2, DOWS, DPOP1, DPOP2, DWRES,
                EVI, LCS, LCT, LCU, LSTD, LSTN,
                MB1, MB2, MB3, MB7, MDEM, NLT, NPPA, NPPS,
                PARA, PARV, PH, S1VV, S2B08, S2B11, S2B12, SLOPE, SND, SOC, TIM, WPOP,
                rainfall_season,
                #meher_rainfall_seas, belg_rainfall_seas, meher_wheat_rainfall_seas,
                #lagrainfall3, lagrainfall6,
                Region) %>% 
  filter(complete.cases(.))

CSA_maize <- CSA %>% 
  filter(!is.na(maize_Unmilled_norm)) %>%
  dplyr::select(maize_Unmilled_norm,
                maize_Unmilled_norm_t0,
                Month_code, Longitude, Latitude,
                BIO1, BIO12, BIO15, BIO7, CEC,
                DCELL, DGRID, DNFPA, DOR1, DOR2, DOWS, DPOP1, DPOP2, DWRES,
                EVI, LCS, LCT, LCU, LSTD, LSTN,
                MB1, MB2, MB3, MB7, MDEM, NLT, NPPA, NPPS,
                PARA, PARV, PH, S1VV, S2B08, S2B11, S2B12, SLOPE, SND, SOC, TIM, WPOP,
                rainfall_season,
                #meher_rainfall_seas, belg_rainfall_seas, meher_wheat_rainfall_seas,
                #lagrainfall3, lagrainfall6,
                Region) %>% 
  filter(complete.cases(.))


CSA_wheat <- CSA %>% 
  filter(!is.na(wheat_unmilled_White_norm)) %>% 
  dplyr::select(wheat_unmilled_White_norm,
                wheat_unmilled_White_norm_t0,
                Month_code, Longitude, Latitude,
                BIO1, BIO12, BIO15, BIO7, CEC,
                DCELL, DGRID, DNFPA, DOR1, DOR2, DOWS, DPOP1, DPOP2, DWRES,
                EVI, LCS, LCT, LCU, LSTD, LSTN,
                MB1, MB2, MB3, MB7, MDEM, NLT, NPPA, NPPS,
                PARA, PARV, PH, S1VV, S2B08, S2B11, S2B12, SLOPE, SND, SOC, TIM, WPOP,
                rainfall_season,
                #meher_rainfall_seas, belg_rainfall_seas, meher_wheat_rainfall_seas,
                #lagrainfall3, lagrainfall6,
                Region) %>% 
  filter(complete.cases(.))

```


# V. EXPORTING THE DATASET

```{r}
write.csv(CSA_barley, "../../data/prices/CSA_barley_market_price_data.csv", row.names=FALSE)
write.csv(CSA_maize, "../../data/prices/CSA_maize_market_price_data.csv", row.names=FALSE)
write.csv(CSA_wheat, "../../data/prices/CSA_wheat_market_price_data.csv", row.names=FALSE)
```
