---
title: "Ensemble Prediction of Market Maize Prices in Ethiopia: Raster surfaces"
author: "Sebastian Palmas"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

# INTRODUCTION
Now that we have the models ready to use to predict prices, we can predict price surfaces for all Ethiopia.

We need to run all ML models and use their predictions for the ensemble method (`model.ens`).

Because the model uses month as one of the predictors and there isn't any raster with month values we need a month to use for the prediction. For that we can use the `const` specification in the `terra::predict` function. We will simulate prices for all months.



This script is the next step from `ETH_maize_market_price_model.Rmd`.


# I. Packages and models
```{r, message=FALSE}
rm(list=ls()) 
library(dplyr)
library(ggplot2)
library(terra)

library(caret)
library(deepnet)
library(gbm)
library(glmnet)
library(pls)
library(randomForest)
```

These are the price models 
```{r, message=FALSE}
#Loading price models
load(file="../../output/models/maip_USDkg_gbm.rda")  #model.gbm
load(file="../../output/models/maip_USDkg_pls.rda")  #model.pls
load(file="../../output/models/maip_USDkg_rf.rda")  #model.rf
load(file="../../output/models/maip_USDkg_ens.rda")  #model.ens
```

# II. Loading rasters
We first need to create a raster stack with all the soil layers.

```{r}
soillist <- list.files(path = "F:/Work/MarkusStacks/ET_250m_2019/ET_500m_2019_resampled/", 
                       pattern = "*.tif$", 
                       full.names = TRUE) # Character vector of relative filepaths

#Creating table of all soil rasterr values
soilrasters <- rast(soillist)

```

In here we load the raster layers for Meher and Belg rainfall seasons that we will use 
```{r}
meher_season <- rast("../../data/rainfall/meher_maize_rainfall_seas_mean_ETH.tif")
belg_season <- rast("../../data/rainfall/belg_rainfall_seas_mean_ETH.tif")

#We also need the region of each pixel to modify the rainfall accordingly.
Meher_Belg_Regions <- rast("../../data/Meher_Belg_Regions/Meher_Belg_Regions.tif")
```

# III. Prediction using a table
```{r}
stack_table <- data.frame(values(soilrasters))

#Let's add this values to the table. These values will be used for the prediction depending on the month and region
stack_table$meher_season <- as.vector(values(meher_season))
stack_table$belg_season <- as.vector(values(belg_season))

#add rainfall to table
stack_table$Meher_Belg_Regions <- as.vector(values(Meher_Belg_Regions))
```

## Adding index and keeping only complete cases
```{r}
#adding index and filtering to complete cases #8,379,171
stack_table <- stack_table %>%
  mutate(index = 1:nrow(.)) %>% 
  filter(complete.cases(.))
```

## Un-trended normalized price value from previous month
The prediction needs a `maize_Unmilled_norm_t0` value which is the un-trended and normalized value of the maize price on the last month.
I'll use a normalized value of 0 to start
```{r}
stack_table$maize_Unmilled_norm_t0 <- 0
```

## Empty Raster to store values
```{r}
blank_raster <- meher_season
values(blank_raster) <- NA
```

## Prediction for all months
```{r}
#Months to predict
Month_codes <- 1
Month_factors <- c(1:8)
for (m in Month_codes){
  #m <- 2 #to test
  #Creating the month code column with factor levels appropriate to use by models
  stack_table$Month_code <- factor(m, levels = Month_factors)
  
  #What happens in the period after belg ends and before Meher ends in regions with Belg rains
  stack_table$rainfall_season <- stack_table$meher_season
  if(m %in% c(6,7)){
    stack_table$rainfall_season[stack_table$Meher_Belg_Regions == 2] <- stack_table$belg_season[stack_table$Meher_Belg_Regions == 2] 
  }
  
  #prediction of all ML methods
  maize_Unmilled_norm.gbm <- predict(model.gbm, newdata=stack_table)
  maize_Unmilled_norm.pls <- predict(model.pls, newdata=stack_table)
  maize_Unmilled_norm.rf <- predict(model.rf, newdata=stack_table)
  
  #Putting together ensemble results to a table
  data.ens <- data.frame(maize_Unmilled_norm.gbm,
                         maize_Unmilled_norm.pls,
                         maize_Unmilled_norm.rf)
  
  #running ensemble model
  maize_Unmilled_norm.ens <- predict(model.ens, data.ens)
  
  #Running ML models
  for(model in c("gbm", "pls", "rf", "ens")){
    values(blank_raster)[stack_table$index] <- get(paste0("maize_Unmilled_norm.", model))
    
    writeRaster(blank_raster, filename=paste0("../../output/crop_market_price/months/ETH_maize_market_price_", model, m, ".tif"),
                overwrite=TRUE, wopt = list(names = paste0("maize_Unmilled_norm.", model)))
    
    plot(blank_raster, main=model)
  }
}
```

# IV. Raster prediction using terra::predict
```{r, eval=FALSE}
#Months to predict
Month_codes <- c(1:8)

#Constant raster to use for maize_Unmilled_norm_t0
maize_Unmilled_norm_t0 <- rast("../../data/rainfall/meher_maize_rainfall_seas_mean_ETH.tif")
values(maize_Unmilled_norm_t0) <- 0
names(maize_Unmilled_norm_t0) <- "maize_Unmilled_norm_t0"

#Constant raster to use for month code
#Month_code <- rast("../../data/rainfall/meher_maize_rainfall_seas_mean_ETH.tif")

#Prediction for all months
for (m in Month_codes){
  #m <- 1 #to tet
  #changing the month to predict
  #values(Month_code) <- m
  #names(Month_code) <- "Month_code"
  
  #constant month factor
  add <- data.frame(Month_code = factor(m, levels=c(1:8)))

  if(m %in% c(6,7)){
    rainfall_season <- rast("../../data/rainfall/meher_maize_rainfall_seas_mean_ETH.tif")
    names(rainfall_season) <- "rainfall_season"
  } else {
    rainfall_season <- rast("../../data/rainfall/meher_maize_rainfall_seas_mean_ETH.tif")
    names(rainfall_season) <- "rainfall_season"
  }
  
  #Putting together the rasterstack with correct rainfall season
  rasterstack <- c(maize_Unmilled_norm_t0, soilrasters, rainfall_season)
  
  #Predicting with ML methods
  maip_USDkg.gbm <- predict(rasterstack, model.gbm, 
                            const = add, na.rm=TRUE,
                            filename=paste0("../../output/crop_market_price/months/ETH_maize_market_price_gbm", m, ".tif"), overwrite=TRUE,
                            wopt = list(names = "maip_USDkg.gbm"))
  maip_USDkg.pls <- predict(rasterstack, model.pls,
                            const = add, , na.rm=TRUE,
                            filename=paste0("../../output/crop_market_price/months/ETH_maize_market_price_pls", m, ".tif"), overwrite=TRUE,
                            wopt = list(names = "maip_USDkg.pls"))
  maip_USDkg.rf <- predict(rasterstack, model.rf, 
                           const = add, na.rm=TRUE,
                           filename=paste0("../../output/crop_market_price/months/ETH_maize_market_price_rf", m, ".tif"), overwrite=TRUE,
                           wopt = list(names = "maip_USDkg.rf"))
  
  #Predicting with ensemble model
  #Putting together rasterstack for ensemble
  #raster.data.ens <- rast(maip_USDkg.pls, 
  #                        maip_USDkg.rf, nlyrs=2)
  
  #maip_USDkg.ens <- predict(raster.data.ens, model.ens, na.rm=TRUE,
  #          filename="../../output/crop_market_price/ETH_maize_market_price_ENS.tif", overwrite=TRUE)

}
```